---
title: "Connecting to the NTSPortal Elasticsearch Search API from R or Python"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Connecting to the NTSPortal Elasticsearch Search API from R or Python}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# API Connection, getting started

The API is called from `elastic.bafg.de`. Users must 1) have been granted API access (please send a request to `ntsportal@bafg.de`) and 2) must have created an API key for themselves at `ntsportal.bafg.de` (Menu->Management->Stack Management->Security->API keys, 'control security privileges' must be checked")

The API documentation is found [here](https://www.elastic.co/docs/api/doc/elasticsearch/). Detailed documentation is available for the different search languages offered, e.g. [Query DSL](https://www.elastic.co/docs/reference/query-languages/querydsl). 

`ntsportal::getTableByEsql()` is the simplest function for collecting data from NTSPortal in R. It uses the `ES|QL` API (See [documentation](https://www.elastic.co/docs/reference/query-languages/esql)) which can be tested in the *Discover* page of the NTSPortal dashboard. The data is returned as a `tbl_df` object (similar to a `data.frame`). The limitation of `ES|QL` is that nested fields (like `ms2` and `compound_annotation`) cannot be queried or retrieved, for this you must use `ntsportal::getTableByQuery()`.

`ntsportal::getTableByQuery()` allows fetching records from NTSPortal as a `tbl_df` and allows the retrieval of nested fields. This function uses the Query DSL language (in the form of a `list`) to pass the queries. The best way to get started with Query DSL is to use the Elasticsearch "Developer Tools" console in the NTSPortal Dashboards. 

For an example of using the API in Python see [Using the Elasticsearch Search API with Query DSL for retrieving documents](articles/Searching-with-ElasticSearch-Query-DSL.html).

## Storing of user credentials

`ntsportal::connectNtsportal()` stores and retrieves the NTSPortal user credentials, it uses `keyring` to store the username and password. 

## Under the hood

The API interface functions above use the Python package *elasticsearch* to query the database ([documentation](https://elasticsearch-py.readthedocs.io/)). To see how it works look at the code for `ntsportal::PythonDbComm()`, which uses the `inst/pythonElasticComm/elasticSearchComm.py` module to create an elasticsearch client object in R. This client is used in many of the interface functions found in `R/dbInterface-PythonDbComm.R`.

### Ignoring SSL verification

This is not recommended, but if there is a certificate error, you can turn off SSL verification when using `elasticsearch.Elasticsearch()` in Python with the argument `verify_certs=False`.

# Checking your IP address

To check the IP from which you are trying to connect:

```{r, eval=FALSE}
library(rjson)
fromJSON(readLines("http://api.hostip.info/get_json.php", warn=F))$ip
```

<!-- Copyright 2025 Bundesanstalt für Gewässerkunde -->
<!-- This file is part of ntsportal -->
