---
title: "Document field descriptions for NTSPortal"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(ntsportal)
```

The following tables contain descriptions for all fields in NTSPortal tables. The dot-notation is used to refer to nested
fields.

# `msrawfiles` tables

These tables hold metadata and processing settings for each raw MS measurement file (mzXML file) in NTSPortal.

```{r, echo=FALSE}
knitr::kable(
  makeMappingsTbl("msrawfiles"),
  caption = "Field descriptions for `msrawfiles` documents",
  align = rep("l", 5)
)
```

# `feature` tables

These table hold processing results ("dbas" and "nts" processing). They are a list of non-target "features" or 
chromatographic peaks with associated metadata.

```{r, echo=FALSE}
knitr::kable(
  makeMappingsTbl("feature"),
  caption = "Field descriptions for `feature` documents",
  align = rep("l", 5)
)
```

# `analysis_dbas` tables

These tables hold results of summary statistics for `feature` tables (currently only for "dbas" processing).

```{r, echo=FALSE}
knitr::kable(
  makeMappingsTbl("analysis_dbas"),
  caption = "Field descriptions for `analysis_dbas` documents",
  align = rep("l", 5)
)
```

# `spectral_library` tables

These documents hold a copy of the [Collective Spectral Library (CSL)](https://doi.org/10.5281/zenodo.16901589).

```{r, echo=FALSE}
knitr::kable(
  makeMappingsTbl("spectral_library"),
  caption = "Field descriptions for `spectral_library` documents",
  align = rep("l", 5)
)
```

# `nondetect_dbas` tables

These documents hold non-detects by file, i.e., which compounds of the CSL were not detected in a measurement file.

```{r, echo=FALSE}
knitr::kable(
  makeMappingsTbl("nondetect_dbas"),
  caption = "Field descriptions for `nondetect_dbas` documents",
  align = rep("l", 5)
)
```
